Casual Bandits

### 2016-01-29 causal-bandit/icml-2015/introduction.tex
Summarise NIPS 2015 paper

Medical drug testing, economic policy setting, and other scientific processes are commonly framed and analysed in the language of sequential experimental design and, in special cases, as bandit problems~\citep{Robbins1952,Chernoff1958}. 
In this framework, single actions or experiments from a pre-determined set are repeatedly performed in order to determine the most effective one via feedback from a single, real-valued reward signal. 
We consider the more general setting for sequential experiments where the feedback after each action consists of the reward signal as well as the values of other variables drawn from a probabilistic causal model~\citep{Pearl2000}.
By explicitly modeling this extra information (\eg, drug side-effects, secondary economic indicators) we show, theoretically and empircally, how it can be used to improve the rate at which effective actions can be identified.

@@

The challenge is best illustrated with an example. Consider a farmer wishing to optimise the yield of her crop. 
She knows that crop yield is only affected by temperature, a particular soil nutrient, and moisture level but the precise effect of their combination is unknown.
In each season the farmer has enough time and money to intervene and control at most one of these variables:
deploying shade or heat lamps will set the temperature to be low or high; the nutrient can be added or removed a through a choice of fertilizer; and irrigation or rain-proof covers will keep the soil wet or dry.
When not intervened upon, the temperature, soil, and moisture vary naturally from season to season due to weather conditions and these are all observed along with the final crop yield at the end of each season.
How might the farmer best experiment to identify the single, highest yielding intervention without sacrificing too much crop yield (relative to always choosing the best intervention) in the process?

This paper can be seen as a first step towards formalising and solving problems such as these. In \S\ref{sec:defs} we introduce \emph{causal bandit problems} in which interventions are treated as arms in a bandit problem but their inlfuence on the reward --- along with any other non-interventional observations --- is assumed to conform to a known causal graph. 
Although the general problem subsumes classical stochastic bandit problems as well as stochastic contextual bandits ... focus on parallel bandit problem.

@@

One obvious approach would be to cast the above example as a stochastic multi-armed bandit problem~\cite{Robbins1952} -- each of the six possible interventions is an arm and the crop yield is the reward -- and apply an algorithm such as UCB~\cite{Auer1995} with well understood regret guarantees.
However, as we show in \S\ref{sec:simple-regret}, this approach ignores the extra information available in the non-intervened variables and subsequently yields a worst-case regret after $T$ round of $\bigomega{\sqrt{NT}}$ with a sub-optimal dependence on $N$, the number of possible interventions.

\todom{Note that feedback work as uses graphs}

Our framework also bears a superficial similarity to contextual bandit problems \cite{Agarwal2014} since the extra observations on non-intervened variables might be viewed as context for selecting an intervention. However, a crucial difference is that in our model the extra observations are only revealed \emph{after} selecting an intevention and cannot be used as context.

There has been a significant amount of recent work which proposes alternative modes of feedback within the bandit setting \cite{TODO} as well as results for partial monitoring which, at first glance, may appear applicable. As we show in Section~\ref{}, applying these results to our problem also exhibit a sub-optimal dependence on the number of available interventions.

@@

An alternate approach to selecting actions is causal inference. Frameworks for causal inference provide a mechanism to specify assumptions that allow observational distributions over variables to be mapped to interventional ones. This allows an agent to predict the outcome of an action based on non-experimental data. This approach is common in social science, demography, and economics where explicit experimentation may be difficult. For example, predicting the effect of changes to childcare subsidies on workforce participation or school choice on student grades.

@@

A natural way to connect the causal framework with the bandit setting is to model the problem as a causal directed acyclic graph. Each possible assignment of variables to values is an action (bandit arm). The reward could be a general function of the action selected and the final state of the graph. However for simplicity, we will consider the reward to be the value of a single specified node minus the cost of the selected action. The number of actions grows exponentially with the number of variables in the graph, making it important to use algorithms that take account of the graph structure to reduce the search space.

Problems framed in this way take on characteristics of different bandit settings depending on the assumptions we make about what subset of actions can be taken, what variables are observable and whether they are observed before or after an action is selected. If feedback is received only on the reward node then the do-calculus can be applied to eliminate some actions immediately, before any experiments are performed and then a standard bandit algorithm can be run on the remaining actions.

If we receive feedback on additional nodes the problem can be more interesting. In addition to being able to eliminate some actions prior to sampling any data as in the previous case, taking one action may give us some information on actions that were not selected.

We consider a bandit problem where the actions and reward are represented by a specific causal graph that demonstrates this interesting structure. We develop an algorithm to leverage the information provided by this structure and demonstrate it substantially outperforms standard bandit algorithms applied to the same problem where the number of actions is large.

@@

Partial monitoring is a very general framework for for decoupling the feedback from the action and reward. It can be used to classify problems into one of four categories, trivial with no regret, easy with $R_T = \bigthetatilde{\sqrt{T}}$ , hard with $R_T = \bigtheta{T^{2/3}}$ and hopeless with $R_T = \bigomega{T}$ \cite{Bartok2014}. Partial monitoring algorithms yield results that are optimal with respect the horizon $T$ but not other parameters, such as $K$, which is the key focus of incorporating causal structure.

ALSO NEED TO MENTION ANY OTHER COMBINATIONS OF BANDITS+CAUSAL (eg the Elias NIPS paper and Generalized Thompson Sampling paper)

Key to Elias' paper is: observing the action an agent would take if it were allowed to make its natural choice can provide some information about hidden confounders that influence both the reward and the choice of action. Therefore, incorporating an agents natural choice as context may outperform a standard bandit that does not use that context. (Note: even in the presence of hidden confounders, including the agents natural choice as context only may improve the results. It is easy to come up with a counter example in which it does not).

>>>>

Medical drug testing, policy setting, and other scientific processes are commonly framed and analysed in the language of sequential experimental design and, in special cases, as bandit problems~\citep{Robbins1952,Chernoff1959}. 
In this framework, single actions (\eg, experiments or interventions) from a pre-determined set are repeatedly performed in order to determine the most effective one via feedback from a single, real-valued reward signal. 
We propose generalizing such sequential experiments by assuming that, in additional to the reward signal, values of other variables are drawn from a probabilistic causal model~\citep{Pearl2000} and observed.
These causal models are common in social science, demography, and economics where explicit experimentation may be difficult. 
For example, predicting the effect of changes to childcare subsidies on workforce participation or school choice on student grades. 
Results from causal inference relate observational distributions to interventional ones, allowing us to predict the outcome of intervention without explicitly performing it. 
By combining causal inference using a causal graph we show, theoretically and empirically, how non-interventional observations can be used to improve the rate at which high-reward actions can be identified.

@@

The type of problem we are concerned with is best illustrated with an example. 
Consider a farmer wishing to optimise the yield of her crop. 
She knows that crop yield is only affected by temperature, a particular soil nutrient, and moisture level but the precise effect of their combination is unknown.
In each season the farmer has enough time and money to intervene and control at most one of these variables:
deploying shade or heat lamps will set the temperature to be low or high; the nutrient can be added or removed a through a choice of fertilizer; and irrigation or rain-proof covers will keep the soil wet or dry.
When not intervened upon, the temperature, soil, and moisture vary naturally from season to season due to weather conditions and these are all observed along with the final crop yield at the end of each season.
How might the farmer best experiment to identify the single, highest yielding intervention without sacrificing too much crop yield (relative to always choosing the best intervention) in the process?

\subsection{Contributions}

This paper can be seen as a first step towards formalising and solving problems such as these. 
In \S\ref{sec:defs} we formally introduce \emph{causal bandit problems} in which interventions are treated as arms in a bandit problem but their influence on the reward --- along with any other observations --- is assumed to conform to a known causal graph. 
We show that our general model subsumes classical stochastic bandit problems as well as stochastic contextual bandits before focusing on a special case where, like the above example, the

@@

One obvious approach would be to cast the above example as a stochastic multi-armed bandit problem (\eg, the six possible farming interventions is an arm and the crop yield is the reward) and apply an algorithm such as UCB~\cite{Auer1995} with well understood regret guarantees.
However, as we show in \S\ref{sec:simple-regret}, this approach ignores the extra information available in the non-intervened variables and subsequently yields a worst-case regret after $T$ round of $\bigomega{\sqrt{NT}}$ with a sub-optimal dependence on $N$, the number of possible interventions.

Our framework also bears a superficial similarity to contextual bandit problems \cite{Langford2008,Agarwal2014} since the extra observations on non-intervened variables might be viewed as context for selecting an intervention. However, a crucial difference is that in our model the extra observations are only revealed \emph{after} selecting an intevention and cannot be used as context.

There has been a significant amount of recent work which proposes alternative modes of feedback within the bandit setting \cite{TODO} as well as results for partial monitoring which, at first glance, may appear applicable. Indeed, we show in Section~\ref{}, that it is possible to applying these existing results to our problem, however the resulting regret is $\bigo{\sqrt{NT}}$, again with a sub-optimal dependence on $N$.

@@

@@

% A natural way to connect the causal framework with the bandit setting is to model the problem as a causal directed acyclic graph. Each possible assignment of variables to values is an action (bandit arm). The reward could be a general function of the action selected and the final state of the graph. However for simplicity, we will consider the reward to be the value of a single specified node minus the cost of the selected action. The number of actions grows exponentially with the number of variables in the graph, making it important to use algorithms that take account of the graph structure to reduce the search space.

% Problems framed in this way take on characteristics of different bandit settings depending on the assumptions we make about what subset of actions can be taken, what variables are observable and whether they are observed before or after an action is selected. If feedback is received only on the reward node then the do-calculus can be applied to eliminate some actions immediately, before any experiments are performed and then a standard bandit algorithm can be run on the remaining actions.

% If we receive feedback on additional nodes the problem can be more interesting. In addition to being able to eliminate some actions prior to sampling any data as in the previous case, taking one action may give us some information on actions that were not selected.

% We consider a bandit problem where the actions and reward are represented by a specific causal graph that demonstrates this interesting structure. We develop an algorithm to leverage the information provided by this structure and demonstrate it substantially outperforms standard bandit algorithms applied to the same problem where the number of actions is large.

@@

% Partial monitoring is a very general framework for for decoupling the feedback from the action and reward. It can be used to classify problems into one of four categories, trivial with no regret, easy with $R_T = \bigthetatilde{\sqrt{T}}$ , hard with $R_T = \bigtheta{T^{2/3}}$ and hopeless with $R_T = \bigomega{T}$ \cite{Bartok2014}. Partial monitoring algorithms yield results that are optimal with respect the horizon $T$ but not other parameters, such as $K$, which is the key focus of incorporating causal structure.

\citet{Bareinboim2015} very recently presented work that also combines bandits and causal inference. Their work differs from our in two key ways. Firstly, their focus is on correcting the effects of unobserved confounding variables whereas we assume all variables are observable. Secondly, they do not assume their learning algorithm has access to a causal graph and instead use a form of contextual randomization to improve reward estimates. Combining their handling of unobserved variables with our analysis is left as future work.

\citet{Ortega2014}

% Key to Elias' paper is: observing the action an agent would take if it were allowed to make its natural choice can provide some information about hidden confounders that influence both the reward and the choice of action. Therefore, incorporating an agents natural choice as context may outperform a standard bandit that does not use that context. (Note: even in the presence of hidden confounders, including the agents natural choice as context only may improve the results. It is easy to come up with a counter example in which it does not).
